On the Programmatic Generation of Reproducible Documents

### 2020-06-23 inst/jss/listdown-jss.tex
And back.

\author{
Michael Kane\\Yale University
}
\title{On the Programmatic Generation of Reproducible Documents}

\Plainauthor{Michael Kane}
\Plaintitle{On the Programmatic Generation of Reproducible Documents}
\Shorttitle{Programmatic Generation of Reproducible Documents}

\Abstract{
The abstract of the article.
}

@@

@@

\begin{CodeInput}
R> doc <- c(
R+   unlist(strsplit(as.character(ld_rmarkdown_header("Anscombe's Quartet",
R+                                author = "Francis Anscombe",
R+                                date = "1973")),
R+                   "\n", fixed = TRUE)),
R+   ld_make_chunks(ld))
R> 
R> doc
\end{CodeInput}

@@

\hypertarget{initialization-code}{%
\section{Initialization code}\label{initialization-code}}

@@

\hypertarget{r-code-chunk-customization}{%
\section{R code chunk customization}\label{r-code-chunk-customization}}

@@

\begin{CodeInput}
R> library(gtsummary)
R> library(dplyr)
R> library(glue)
R> library(survival)
R> library(survminer)
R> library(rmarkdown)
R> 
R> # Function to create the computational components.
R> make_surv_cc <- function(trial, treat, surv_cond_chars) {
R+   # Create Table 1 by treatment
R+   table_1 <- trial %>%
R+     tbl_summary(by = all_of(treat)) %>%
R+     gtsummary::as_flextable()
R+ 
R+   # Create survival plots base on survival characteristics
R+   scs <- lapply(c("1", surv_cond_chars),
R+                 function(sc) {
R+                   glue("Surv(ttdeath, death) ~ {treat} + {sc}") %>%
R+                     as.formula() %>%
R+                     surv_fit(trial) %>%
R+                     ggsurvplot()
R+                 })
R+   names(scs) <- c("Overall", tools::toTitleCase(surv_cond_chars))
R+   list(`Table 1` = table_1, `Survival Plots` = scs)
R+ }
R> 
R> # Create the computational components and write to surv-cc.rds.
R> surv_cc <- make_surv_cc(trial, treat = "trt",
R+                         surv_cond_chars = c("stage", "grade"))
R> 
R> ld_cc_dendro(surv_cc)
\end{CodeInput}

@@

\begin{CodeInput}
R> saveRDS(surv_cc, "surv-cc.rds")
R> 
R> # Create the listdown object with appropriate dependencies
R> # and presentation parameters.
R> ld_surv <- listdown(load_cc_expr = readRDS("surv-cc.rds"),
R>                     package = c("gtsummary", "survminer", "flextable"),
R>                     decorator = list(ggsurvplot = identity,
R>                                      tbl_summary = identity),
R>                     echo = FALSE,
R>                     message = FALSE,
R>                     warning = FALSE,
R>                     fig.width = 7,
R>                     fig.height = 4.5)
R> 
R> # Write an RMarkdown header and chunks derived from the 
R> # computational components and listdown object to trial-report.rmd.
R> writeLines(
R>   paste(c(
R>     as.character(ld_rmarkdown_header("Simple Trial Report")),
R>     ld_make_chunks(ld_surv))),
R>   "trial-report.rmd")
R> 
R> # Render the RMarkdown file to html
R> render("trial-report.rmd", quiet = TRUE, html_document(keep_tex = TRUE))  
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> ---
R> title: Simple Trial Report
R> output: html_document
R> ---
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
R> library(gtsummary)
R> library(survminer)
R> library(flextable)
R> 
R> cc_list <- readRDS("surv-cc.rds")
R> \```
R> 
R> # Table 1
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
R> cc_list$`Table 1`
R> \```
R> 
R> # Survival Plots
R> 
R> ## Overall
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
R> cc_list$`Survival Plots`$Overall
R> \```
R> 
R> ## Stage
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
R> cc_list$`Survival Plots`$Stage
R> \```
R> 
R> ## Grade
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
R> cc_list$`Survival Plots`$Grade
R> \```
\end{CodeInput}
\end{CodeChunk}

>>>>

\author{
Michael Kane\\Yale University \And Simon Urbanek\\The University of
Aukland
}
\title{On the Programmatic Generation of Reproducible Documents}

\Plainauthor{Michael Kane, Simon Urbanek}
\Plaintitle{On the Programmatic Generation of Reproducible Documents}
\Shorttitle{On the Programmatic Generation of Reproducible Documents}

\Abstract{
Reproducible document standards, like \proglang{R} Markdown, facilitate
the programmatic creation of documents whose content is itself
programmatically generated. While these documents are generally not
complete in the sense that they will not include prose content,
generated by an author to provide context, a narrative, etc.,
programmatic generation can provide substantial efficiencies for
structuring and constructing documents. This paper explores the
programmatic generation of reproducible by distinguishing components
than can be created by computational means from those requiring
human-generated prose, providing guidelines for the generation of these
documents, and identifying a use case in clinical trial reporting. These
concepts and use case are illustrated through the \pkg{listdown} package
for the \proglang{R} programming environment, which is is currently
available on the Comprehensive R Archive Network (CRAN).
}

@@

      Simon Urbanek\\
  The University of Aukland\\
  38 Princes StreetAuckland Central, Aukland 1010, NZ\\
  E-mail: \href{mailto:s.urbanek@auckland.ac.nz}{\nolinkurl{s.urbanek@auckland.ac.nz}}\\

@@

\begin{CodeInput}
R> doc <- c(
R+   as.character(ld_rmarkdown_header("Anscombe's Quartet",
R+                                    author = "Francis Anscombe",
R+                                    date = "1973")),
R+   ld_make_chunks(ld))
R> 
R> doc
\end{CodeInput}

@@

\hypertarget{advanced-usage}{%
\section{Advanced usage}\label{advanced-usage}}

\hypertarget{initialization-code}{%
\subsection{Initialization code}\label{initialization-code}}

@@

\hypertarget{r-code-chunk-customization}{%
\subsection{R code chunk
customization}\label{r-code-chunk-customization}}

@@

\begin{CodeInput}
R> library(gtsummary)
R> library(dplyr)
R> library(survival)
R> library(survminer)
R> library(rmarkdown)
R> 
R> # Function to create the computational components.
R> make_surv_cc <- function(trial, treat, surv_cond_chars) {
R+   # Create Table 1 by treatment
R+   table_1 <- trial %>%
R+     tbl_summary(by = all_of(treat)) %>%
R+     gtsummary::as_flextable()
R+ 
R+   # Create survival plots base on survival characteristics
R+   scs <- lapply(c("1", surv_cond_chars),
R+                 function(sc) {
R+                   sprintf("Surv(ttdeath, death) ~ %s + %s", treat, sc) %>%
R+                     as.formula() %>%
R+                     surv_fit(trial) %>%
R+                     ggsurvplot()
R+                 })
R+   names(scs) <- c("Overall", tools::toTitleCase(surv_cond_chars))
R+   list(`Table 1` = table_1, `Survival Plots` = scs)
R+ }
R> 
R> # Create the computational components and write to surv-cc.rds.
R> surv_cc <- make_surv_cc(trial, treat = "trt",
R+                         surv_cond_chars = c("stage", "grade"))
R> 
R> ld_cc_dendro(surv_cc)
\end{CodeInput}

@@

\begin{CodeInput}
R> # Turn the survplots into a list to show the individual components.
R> class(surv_cc$`Survival Plots`$Overall) <- 
R+   class(surv_cc$`Survival Plots`$Stage) <-
R+   class(surv_cc$`Survival Plots`$Grade) <- "list"
R> 
R> # Create a tab for each element of the survplot.
R> names(surv_cc$`Survival Plots`) <- 
R+   paste(names(surv_cc$`Survival Plots`), "{.tabset}")
R> 
R> # Rename the elements
R> names(surv_cc$`Survival Plots`$`Overall {.tabset}`) <- 
R+   names(surv_cc$`Survival Plots`$`Stage {.tabset}`) <- 
R+   names(surv_cc$`Survival Plots`$`Grade {.tabset}`) <- 
R+   c("Plot", "Data", "Table")
R>   
R> # Save the updated object for presentation.
R> saveRDS(surv_cc, "surv-cc.rds")
R> 
R> # Create the listdown object with appropriate dependencies
R> # and presentation parameters.
R> ld_surv <- listdown(load_cc_expr = readRDS("surv-cc.rds"),
R+                     package = c("gtsummary", "flextable", "DT", 
R+                                 "ggplot2"),
R+                     decorator_chunk_opts = 
R+                       list(gg = list(fig.width = 8,
R+                                      fig.height = 6)),
R+                     decorator = list(data.frame = datatable),
R+                     echo = FALSE,
R+                     message = FALSE,
R+                     warning = FALSE,
R+                     fig.width = 7,
R+                     fig.height = 4.5)
R> 
R> # Write an RMarkdown header and chunks derived from the 
R> # computational components and listdown object to trial-report.rmd.
R> writeLines(
R+   paste(c(
R+     as.character(ld_rmarkdown_header("Simple Trial Report")),
R+     ld_make_chunks(ld_surv))),
R+   "trial-report.rmd")
R> 
R> # Render the RMarkdown file to html and show it.
R> render("trial-report.rmd", quiet = TRUE)
R> browseURL("trial-report.html")
\end{CodeInput}
\end{CodeChunk}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

While the programmtic generation of reproducible documents has appealing
qualities, it also has fundamental limitations that should be kept in
mind when deciding if tools like \pkg{listdown} should be employed.
First and foremost, without narrative components a document has very
little context. Quantitative analyses require research questions,
hypotheses, reviews, interpretations, and conclusions. Computational
components are necessary but generally not sufficient for constructing
an analysis. This means that if narrative components must be conveyed in
a document, then \pkg{listdown} may make the generation of their
presentation more convenient. However, it does not relieve the burden of
the author to create prose developing a narrative.

Second, it is difficult if not impossible to construct computational
components for an arbitary analyses. Analyses themselves have context
and are built with a set of assumptions and goals. Our experience shows
that \pkg{listdown} is easiest to used for a fixed data format. This
means a standard set of table and visualizations for similarly formatted
data. In particular, data that is periodicaly updated, without changes
to the format, is a case that is particularly amenable to document
generation reuse.

Keeping these limitations in mind, \pkg{listdown} can be used to
effectively reduce the the difficulty of generating documents in a
variety of contexts and fits readily in data processing and analysis
pipelines.

@@

\begin{CodeInput}
R> ---
R> title: Simple Trial Report
R> output: html_document
R> ---
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> library(gtsummary)
R> library(flextable)
R> library(DT)
R> library(ggplot2)
R> 
R> cc_list <- readRDS("surv-cc.rds")
R> \```
R> 
R> # Table 1
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> cc_list$`Table 1`
R> \```
R> 
R> # Survival Plots
R> 
R> ## Overall {.tabset}
R> 
R> ### Plot
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, 
R>      fig.height = 6}
R> cc_list$`Survival Plots`$`Overall {.tabset}`$Plot
R> \```
R> 
R> ### Data
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> datatable(cc_list$`Survival Plots`$`Overall {.tabset}`$Data)
R> \```
R> 
R> ### Table
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> datatable(cc_list$`Survival Plots`$`Overall {.tabset}`$Table)
R> \```
R> 
R> ## Stage {.tabset}
R> 
R> ### Plot
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, 
R>      fig.height = 6}
R> cc_list$`Survival Plots`$`Stage {.tabset}`$Plot
R> \```
R> 
R> ### Data
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> datatable(cc_list$`Survival Plots`$`Stage {.tabset}`$Data)
R> \```
R> 
R> ### Table
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> datatable(cc_list$`Survival Plots`$`Stage {.tabset}`$Table)
R> \```
R> 
R> ## Grade {.tabset}
R> 
R> ### Plot
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, 
R>      fig.height = 6}
R> cc_list$`Survival Plots`$`Grade {.tabset}`$Plot
R> \```
R> 
R> ### Data
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> datatable(cc_list$`Survival Plots`$`Grade {.tabset}`$Data)
R> \```
R> 
R> ### Table
R> 
R> \```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, 
R>      fig.height = 4.5}
R> datatable(cc_list$`Survival Plots`$`Grade {.tabset}`$Table)
R> \```
\end{CodeInput}
\end{CodeChunk}
